\documentclass[11pt, a4paper]{article}
\usepackage{amsmath, amssymb, hyperref, geometry, booktabs}
\geometry{total={170mm,257mm}, left=25mm, top=25mm}
\title{\textbf{ESCCML Study: C2. Codec and Canonical Determinism}}
\author{ESCCML Technical Working Group}
\date{Version 1.0}

\begin{document}
\maketitle

\begin{abstract}
This study defines the Codec primitive of ESCCML (Epoch-Split Codec Convergence Manifest Ledger).
The Codec establishes byte-for-byte deterministic serialization across all architectures, ensuring
auditability, replayability, and compression efficiency. By separating high-salience value data from
lower-salience metadata, the Codec enables zero-copy Merkle hashing, lightweight verification, and
energy-efficient state propagation.
\end{abstract}

\section{The Codec: Packed-Byte Channelized Encoding}
The \textbf{Codec} is the ESCCML standard for data serialization, designed to ensure
\textbf{deterministic replayability} across heterogeneous hardware and to maximize compression
without compromising integrity.

\subsection{Channelized Encoding Model (Y/Cb/Cr Analogy)}
The Codec utilizes a channelized encoding structure, conceptually analogous to the Y/Cb/Cr color
model, to separate data by salience and compression priority:
\begin{itemize}
    \item \textbf{Y-Channel (Luminance $\rightarrow$ Value):} Carries high-salience, canonical data,
    such as balances, nonces, and transfer amounts. This channel is always fixed-width and
    high-precision, ensuring integrity and enabling zero-copy Merkle hashing.
    \item \textbf{Cb/Cr Channels (Chrominance $\rightarrow$ Metadata):} Handles lower-salience,
    compressible data, such as optional proofs, public keys, and human-readable metadata.
    These channels support variable-length encoding, palette dictionaries, and LZ4-compressed gossip.
\end{itemize}

This separation allows \textbf{light verification clients} to process only the Y-channel, while
full nodes can decode all channels for complete state reconstruction.

\section{Mandates for Canonical Determinism}
To achieve post-blockchain auditability, the Codec enforces strict rules against
non-deterministic artifacts.

\subsection{Strict Integer-Only Accounting}
All value representation in ESCCML must be handled using $\mathbf{integer-only\ arithmetic}$.
This eliminates the potential for $\mathbf{floating-point\ drift}$ errors and rounding
discrepancies in financial state reconciliation, a common failure point in legacy ledgers.

\subsection{Codec Version Hash and Integrity}
The Manifest must cryptographically bind to its Codec specification:
\begin{itemize}
    \item The $\mathbf{Codec\ Version\ Hash}$ is embedded in the Manifest digest.
    \item ESCCML Validators are mandated to $\mathbf{reject\ mixed-version\ replays}$.
    Any state proposed using a Codec version different from the anchored state must be flagged
    and rejected by the convergence model, guaranteeing all nodes interpret the state identically.
\end{itemize}

\section{Compression and Efficiency Considerations}
The Codec is designed to minimize I/O and CPU cycles:
\begin{itemize}
    \item Palette-based address encoding reduces redundancy in manifests.
    \item Huffman-ready tables allow deterministic compression without entropy drift.
    \item Subsampling of metadata channels ensures bandwidth efficiency for gossip propagation.
\end{itemize}
Benchmarks from Phase 0 validation demonstrate that Codec-native processing contributes
significantly to the ultra-low energy profile of ESCCML ($\sim 1.4 \times 10^{-12}$ kWh/tx).

\section{Implications for Light Clients and Interop}
\begin{itemize}
    \item Light clients can verify balances and transfers by decoding only the Y-channel.
    \item Cross-network interop is simplified by embedding Codec hashes in manifests,
    ensuring deterministic replay across heterogeneous networks.
    \item Future confidential transaction envelopes will extend the Codec without breaking
    determinism, by versioning and embedding proof commitments in metadata channels.
\end{itemize}

\section{Conclusion}
The Codec is the backbone of ESCCML determinism. By separating value from metadata,
enforcing integer-only arithmetic, and binding manifests to Codec versions, ESCCML ensures
that every node, regardless of hardware or geography, interprets state transitions identically.
This guarantees forensic auditability, lightweight verification, and energy-efficient operation.

\end{document}
